---
title: "Assignment 3"
author: "Divya Singireddy"
date: "2024-03-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load required libraries

```{r}

library(tidyverse)

```
```{r}

library(reshape)

```

```{r}

library(reshape2)

```

```{r}

library(caret)

```

```{r}

library(e1071)

```


##Load the data csv file

```{r}

Universal_Bank <- read.csv("C:/Users/divya/Downloads/UniversalBank (1).csv")

head(Universal_Bank)

```

```{r}

tail(Universal_Bank)

```

#Defining Column names
```{r}

colnames(Universal_Bank)

```

#Data transformation into factors and categories.

```{r}

Universal_Bank$`Personal Loan` = as.factor(Universal_Bank$Personal.Loan)
Universal_Bank$Online = as.factor(Universal_Bank$Online)
Universal_Bank$CreditCard = as.factor(Universal_Bank$CreditCard)

```

#Dividing the data in half, with 40% going into the validation set and 60% going into the training set

```{r}
set.seed(456)

```


#60 % training data and 40% Validation data respectively

```{r}

UniBank.train.data <- sample(row.names(Universal_Bank), 0.6*dim(Universal_Bank)[1])
UniBank.valid.data <- setdiff(row.names(Universal_Bank), UniBank.train.data)

```

#Putting training and validation data into separate dataframes from the dataset

```{r}

UniBank.train <- Universal_Bank[UniBank.train.data, ]
UniBank.valid <- Universal_Bank[UniBank.valid.data, ]

```

#Duplicating the data frame UB.train

```{r}

train <- Universal_Bank[UniBank.train.data, ]

```

#Duplicating the data frame UB.valid

```{r}

valid <- Universal_Bank[UniBank.train.data,]

```

## A.Create a pivot table for the training data with Online as a column variable, CC as a row variable, and Loan as a secondary row variable. The values inside the table should convey the count. In R use functions melt() and cast(), or function table(). In Python, use panda dataframe methods melt() and pivot().

```{r}

## melt data from data
melted_data = melt(train,id=c("CreditCard","Personal.Loan"),variable= "Online")

```

```{r}

# Casting the melted data
cast_data <- dcast(melted_data, CreditCard + Personal.Loan ~ value, fun.aggregate = length)
# Casting column number 14: Personal loan, ID, and credit card, respectively
cast_data[,c(1,2,3,14)]

```

## B. Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)].

```{r}

#According to the pivot table, the value for Loan is 89, and the value for CC is
UniversalBank.Loan.CC1 <- 89/3000

# which is 29 %.
UniversalBank.Loan.CC1

```

## C. Create two separate pivot tables for the training data. One will have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC.

```{r}

## reshaping the train data frame into a long format, with "Personal.Loan" as an identifier and "Online" as a variable to be melted
melt1 = melt(train,id=c("Personal.Loan"),variable = "Online")

```

```{r}

## reshaping the train data frame into a long format, with "CreditCard" as an identifier and "Online" as a variable to be melted
melt2 = melt(train,id=c("CreditCard"),variable = "Online")

```

```{r}

# Casting Personal loan and online values
# Assuming both Personal.Loan and Online are binary (0 or 1)
cast1 =dcast(melt1,`Personal.Loan`~Online)

```

```{r}

# Casting Personal loan and online values
cast2=dcast(melt2,CreditCard~Online)

```

```{r}

#shows the number of personal loans in reference to online
UniversalBank.Loanonline=cast1[,c(1,13)]
UniversalBank.LoanCC = cast2[,c(1,14)]
UniversalBank.Loanonline

```

```{r}

# shows the number of credit cards in reference to Online.
UniversalBank.LoanCC

```

## D. Compute the following quantities [P (A | B) means “the probability of A given B”]: #1.P (CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors) #2.P(Online=1|Loan=1) #3.P (Loan = 1) (the proportion of loan acceptors) #4.P(CC=1|Loan=0) #5.P(Online=1|Loan=0) #6.P(Loan=0)

```{r}

# creating a pivot table with the columns 14 and 10 representing personal loan
table(train[,c(14,10)])

```

```{r}

# Creating a pivot table for column 13 and 10 which is online and personal loan
table(train[,c(13,10)])

```

```{r}

# Personal loan pivot table There are 2725 and 275 from training, respectively
table(train[,c(10)])

```

#1.P (CC = 1 | Loan = 1) 

```{r}

## We can obtain the CC= 1 and Loan = 1 values by referring to the above p
UniversalBank.CCUniversalBank.Loan1 = 89/(89+200)
UniversalBank.CCUniversalBank.Loan1

```

#2.P(Online=1|Loan=1)

```{r}

#We can get the online = 1 and loan = 1 values from the pivot table aboveUB.ONUB.Loan1
UniversalBank.ONUniversalBank.Loan1 =177/(177+112)

```


#3.P (Loan = 1)

```{r}

# By referring the above pivot table we can get the Loan = 1
UniversalBank.Loan1 =289/(289+2711)
UniversalBank.Loan1

```

#4.P(CC=1|Loan=0) 

```{r}

#Using the pivot table above, we can obtain the CC = 1 and Loan = 0 values
Universal_Bank.CCLoan.01= 794/(794+1917)
Universal_Bank.CCLoan.01

```
#5.P(Online=1|Loan=0)

```{r}

# We can get the online = 1 and loan = 0 values from the pivot table above.
Universal_Bank.ON1.L0= 1665/(1665+1046)
Universal_Bank.ON1.L0

```

#6.P(Loan=0)

```{r}

# We can obtain the Loan = 0 values by the pivot table above.
Universal_Bank.Loan0= 2711/(2711+289)
Universal_Bank.Loan0

```

## E. Use the quantities computed above to compute the naive Ba1 probability P(Loan = 1 | CC = 1,Online = 1).

```{r}

# Given probabilities
P_CC_given_Loan1 <- 0.096
P_Online_given_Loan1 <- 0.833
P_Loan1 <- 0.0125

```


```{r}

# Calculate Naive Bayes probability P(Loan = 1 | CC = 1, Online = 1)
#Universal_Bank_Naivebayes <- (P_CC_given_Loan1 * P_Online_given_Loan1 * P_Loan1) / (P_CC * P_Online)
Universal_Bank_Naivebayes <- (P_CC_given_Loan1)*(P_Online_given_Loan1)*(P_Loan1)
Universal_Bank_Naivebayes

```

## F. Compare this value with the one obtained from the pivot table in (b). Whichis  a more accurate estimate?

```{r}

# Naive Bayes Probability (from calculation in E)
naive_bayes_probability <- 0.0009996
# Pivot Table Probability (from part B)
pivot_table_probability <- 0.02966667
# Compare the probabilities and print a message indicating which is more accurate
if (naive_bayes_probability > pivot_table_probability) {
message("Naive Bayes Probability is more accurate: ", naive_bayes_probability)
} else if (naive_bayes_probability < pivot_table_probability) {
message("Pivot Table Probability is more accurate: ", pivot_table_probability)
} else {
message("Both probabilities are the same: ", naive_bayes_probability)
}

```

Pivot Table Probability is more accurate: 0.02966667
Based on the comparison, the pivot table probability (0.02966667) is considered more accurate compared to the Naive Bayes probability.

## G. Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).

```{r}

names(Universal_Bank)

```

```{r}

names(UniBank.train)

```

```{r}

# Select the relevant columns for training
UniBank.train<- Universal_Bank[, c("CreditCard", "Online", "Personal Loan")]

# Rename the columns to remove spaces
colnames(UniBank.train)<- c("CreditCard", "Online", "PersonalLoan")

# Convert 'CreditCard' and 'Online' to factors with appropriate levels
UniBank.train$CreditCard <- factor(UniBank.train$CreditCard, levels = c(0, 1), labels = c("No", "Yes"))

UniBank.train$Online <- factor(UniBank.train$Online, levels = c(0, 1), labels = c("No", "Yes"))

```


```{r}

# Print the probability
print("Probability of loan = 1 given CC = 1 and Online = 1:")

```








